---
title: "Bellabeat"
author: "Karishma"
date: "2023-12-20"
output:
  pdf_document: default
  html_document: default
---
## Introduction 

Bellabeat is a high-tech manufacturer of health-focused products for women. Collecting data on activity, sleep, stress, and reproductive health has allowed Bellabeat to empower women with knowledge about their own health and habits. Although Bellabeat is a successful small company, they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company.

## 1. Ask

##### Business Task: 
To identify potential opportunities for growth and provide recommendations for the Bellabeat marketing strategy improvement based on trends in smart device usage.

##### Key Stakeholders:
* Urška Sršen: Bellabeat's cofounder and Chief Creative Officer
* Sando Mur: Mathematician and Bellabeat’s co-founder

##### Questions to explore for the analysis:
* What are some trends in smart device usage?
* How could these trends apply to Bellabeat customers?
* How could these trends help influence Bellabeat marketing strategy?

## 2. Prepare

This Kaggle data set contains personal fitness tracker from thirty fitbit users. Thirty eligible Fitbit users consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring. It includes information about daily activity, steps, and heart rate that can be used to explore users’ habits.
 
#### Loading Packages
```{r}
library(tidyverse)
library(lubridate) 
library(dplyr)
library(ggplot2)
library(tidyr)
library(janitor)
```
## 3. Process

#### Importing the Datasets
```{r}
# Read the dataframes
activity <- read_csv("C:/Users/karis/Downloads/input/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
calories <- read_csv("C:/Users/karis/Downloads/input/Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv")
intensities <- read_csv("C:/Users/karis/Downloads/input/Fitabase Data 4.12.16-5.12.16/hourlyIntensities_merged.csv")
sleep <- read_csv("C:/Users/karis/Downloads/input/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
weight <- read_csv("C:/Users/karis/Downloads/input/Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
```
#### data
```{r}
head(activity)
```
```{r}
colnames(activity)
```
```{r}
head(weight)
```
```{r}
colnames(weight)
```
#### Converting date time format
```{r}
# intensities
intensities$ActivityHour=as.POSIXct(intensities$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
intensities$time <- format(intensities$ActivityHour, format = "%H:%M:%S")
intensities$date <- format(intensities$ActivityHour, format = "%m/%d/%y")
# activity
activity$ActivityDate=as.POSIXct(activity$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
activity$date <- format(activity$ActivityDate, format = "%m/%d/%y")
# sleep
sleep$SleepDay=as.POSIXct(sleep$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
sleep$date <- format(sleep$SleepDay, format = "%m/%d/%y")
```

## 4. Analyze
#### Number of Participants in each category
```{r}
n_distinct(activity$Id)  
n_distinct(calories$Id)   
n_distinct(intensities$Id)
n_distinct(sleep$Id)
n_distinct(weight$Id)
```
To summarize the above data, there are 33 participants in the activity, calories, and intensities datasets, 24 in the sleep dataset, and only 8 in the weight dataset. The fact that there are only 8 participants in the weight dataset means that more data would be needed to make a strong reccomendation or conclusion.

#### checking for significant change in weight
```{r}
weight%>%
 group_by(Id)%>%
  summarise(min(WeightKg),max(WeightKg))
```
There is no significant changes in weight of 8 participants. 

#### The summaries for the rest of the datasets:
```{r}
# activity
activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes, Calories) %>%
  summary()

# active minutes per category
activity %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes) %>%
  summary()

# calories
calories %>%
  select(Calories) %>%
  summary()
# sleep
sleep %>%
  select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>%
  summary()
# weight
weight %>%
  select(WeightKg, BMI) %>%
  summary()
```
#### Observations made from the above summaries:
* Sedetary minutes on average is 16.5 hours.
* The average number of steps per day is 7638. The CDC recommends people take 10,000 steps daily.
* The majority of the participants are lightly active.
* The average participant burns 97 calories per hour.
* On an average, participants sleep for 7 hours.

#### Merging Data
Merging two datasets Activity and Sleep on Columns Id and date.

```{r}
merged_data <- merge(sleep, activity, by = c('Id', 'date'))
head(merged_data) 
```
## 5. Share

```{r}
ggplot(data = activity, aes(x = TotalSteps, y = Calories)) + geom_point() + geom_smooth() + labs(title = "Total Steps vs. Calories")
```
There is a correlation between total number of steps taken and calories burned. The more steps each participant takes, the more calories they burn.

```{r}
ggplot(data = sleep, aes(x = TotalMinutesAsleep, y = TotalTimeInBed)) + geom_point() + labs(title = "Total time asleep vs Total time in bed")
```
There is a positive correlation between total time asleep vs total time in bed. To improve sleep quality for its users, bellabeat should consider having a section where users can customize their sleep schedule to ensure consistency.

```{r}
ggplot(data = merged_data, mapping = aes(x = SedentaryMinutes, y = TotalMinutesAsleep)) + 
  geom_point() + labs(title= "Sleep Duration and Sedentary Time")
```
```{r}
cor(merged_data$TotalMinutesAsleep,merged_data$SedentaryMinutes)
```
There is a negative correlation between SedentaryMinutes and TotalMinutesAsleep. This means that the less active a participant is, the less sleep they tend to get.

#### Whether the day of the week affects our activity levels and sleep.
```{r}
# aggregate data by day of week to summarize averages 
merged_data <- mutate(merged_data,day = wday(SleepDay, label = TRUE))
summarized_activity_sleep <- merged_data %>% 
  group_by(day) %>% 
  summarise(AvgDailySteps = mean(TotalSteps),
            AvgAsleepMinutes = mean(TotalMinutesAsleep),
            AvgAwakeTimeInBed = mean(TotalTimeInBed), 
            AvgSedentaryMinutes = mean(SedentaryMinutes),
            AvgLightlyActiveMinutes = mean(LightlyActiveMinutes),
            AvgFairlyActiveMinutes = mean(FairlyActiveMinutes),
            AvgVeryActiveMinutes = mean(VeryActiveMinutes), 
            AvgCalories = mean(Calories))
head(summarized_activity_sleep)
```
```{r}
ggplot(data = summarized_activity_sleep, mapping = aes(x = day, y = AvgDailySteps)) +
geom_col(fill = "green") + labs(title = "Daily Step Count")
```
The bar graph above shows us that participants are most active on saturdays and least active on sundays.

## 6. Act

After analyzing the FitBit Fitness Tracker data, I came up with some recommendations for Bellabeat marketing strategy based on trends in smart device usage.

* The majority of participants are lightly active. Bellabeat should offer a progression system in the app to encourage participants to become at least fairly active.
* If users want to improve the quality of their sleep, Bellabeat should consider using app notifications reminding users to get enough rest, as well as recommending reducing sedentary time.
* Participants are most active on Saturdays. Bellabeat can use this knowledge to remind users to go for a walk or a jog on that day. Participants seem to be the least active on Sundays. Bellabeat can use this to motivate users to go out and continue exercising on Sundays.
